The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-25 paper.tex
Completed rewrite of introduction

@@

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time. In robotics, anthropomorphism is traditionally
perceived as a static feature triggered by the human-like design of a robot that once observed (during a
short-term interaction) reflects a sustaining social effect. Based on a previously published
literature review~\citep{fink_anthropomorphism_2012}, a
long-term field study in a natural environment~\citep{fink_living_2013}, a 
content analysis of online discussion 
forums~\citep{fink_anthropomorphic_2012}, and an on-going child-robot 
interaction study, we believe that anthropomorphic
effects are based on several factors, and that they do evolve over time in 
non-monotonic ways. Moreover, we assume that these variances and dynamics 
play a central role in sustaining engagement in human-robot interaction.

This article is organized around two main hypothesis that we formulated based on 
our empirical studies of HRI, and by integrating previous work from several domains.

\begin{itemize}
    \item \textbf{Hypothesis 1~} A person's tendency to anthropomorphize a robot
is based on several factors, including (but not limited to) (1) the 
design of the robot, (2) the personal characteristics of the human user, 
and (3) the context and situation of use. For instance, a social 
context and purpose of the robot increase anthropomorphism. 
\textit{(anthropomorphism as a multi-factor phenomenon)}
	\item \textbf{Hypothesis 2~} A person's tendency to anthropomorphize a robot 
evolves in a non-linear way over time and with growing experience that the 
person has with the robot. For instance, a more experienced user will perceive 
the robot as more familiar and predictable, and in turn anthropomorphizes the 
robot less. \textit{(anthropomorphism as a dynamic phenomenon)}
\end{itemize}

We propose to formalize these variances and dynamic effects into a model that we call the
\emph{Dynamics of Anthropomorphism}. We hope that it may support further
discussions and reflections on how anthropomorphism impacts human-robot
interaction on the long run, and also foster research on the affective bonds
induced by anthropomorphic projections on robots.
Through our model, we propose to represent how the level of anthropomorphic effects (\ie
observable manifestations of anthropomorphism) evolves over a long-term
human-robot interaction (Figure~\ref{fig:dynamics}). By long-term interaction,
we mean direct (non-mediated), repeated interaction with the same robot, over
an extended period of time (typically longer than a week).

%In this article, we would further like to share some critical but constructive
%considerations on anthropomorphism in robotics and HRI. Thereby, we contribute
%to a better understanding of this complex phenomenon related to robots.

The remainder of this article is organized as follows.  First, in Section
\ref{sec:overview-studies}, we give a brief overview on the various studies that
we conducted to investigate anthropomorphism, namely (1) the Roomba study, (2)
the Forum study, (3) the Ranger domino study. Next, in Section 
\ref{sec:anthropomorphism} we aim to establish a common ground in the understanding 
of anthropomorphism. We first explore an artifact-centered perspective and briefly
outline how the human-like design of agents can encourage anthropomorphism.
Then, we explore a human-centered perspective by integrating theories from
cognitive science, developmental and social psychology that serve to explain
anthropomorphism. At the end of this section we bring both perspectives together
and interpret that anthropomorphism is a multi-factor phenomenon.  Based on the
two previous sections, we propose an initial model of the \textit{``Dynamics of
Anthropomorphism''} in Section \ref{sec:our-ideas}. This model describes the
variances and dynamics of people's tendency to anthropomorphize an artificial
agent, such as a robot.  Next, in Section \ref{sec:cognition-neuroscience} we
present a cognitive and neuroscience perspective on anthropomorphism to
consolidate our considerations and the proposed model.  This will be followed by
a discussion about the limitations of our approach (Section
\ref{sec:discussion}). Further, we debate existing and alternative ways to
measure anthropomorphism in human-robot interactions.  Finally, Section
\ref{sec:conclusion} concludes this article and offers starting points for
further research on anthropomorphism and the effect of the human-like design
approach in robotics.

>>>>

\usepackage{paralist} % inparaenum

@@

These first remarks lead to the main question we attempt to address in this
article: how can we understand \emph{anthropomorphism} as a whole, and in such a
way that is support improved human-robot interactions?

Based on an extensive literature review~\citep{fink_anthropomorphism_2012},
several field and lab studies~\citep{ fink_anthropomorphic_2012,
fink_living_2013, fink2014which, lemaignan2014dynamics} and insights from the
cognitive neurosciences, we propose to discuss anthropomorphism from three
complementary perspectives.

First, we would like to clarify what anthropomorphism actually \emph{means}:
besides the \emph{artifact-centered} understanding of anthropomorphism, it
seems important to explicit the human and social factors of anthropomorphism.

Then, we propose a model to understand how anthropomorphism evolves over time.
These dynamics are non-monotonic, and psychological factors like the
\emph{novelty effect} play here a key role.

Finally, we adopt a third perspective on anthropomorphism by investigating its
\emph{cognitive correlates}. Several studies previously showed that we build
instinctive bonds with robots as soon as we attribute animacy, and we also know
from social psychology some of the mechanisms that lead to
\emph{familiarization}. We attempt here to ``connect the dots'', in order to
build a practical model of the cognitive stages we encounter in human-robot
interaction.

\paragraph{Two hypotheses} Our research question also translates to two
hypotheses, which structure the reflections of this article:

\item A person's tendency to anthropomorphize a robot
    is based on several factors, including (but not limited to) (1) the
    design of the robot, (2) the personal characteristics of the human user,
    and (3) the context and situation of use. For instance, a social context
    and purpose of the robot increase anthropomorphism: \textit{anthropomorphism
    as a multi-factor phenomenon}.

\item A person's tendency to anthropomorphize a robot
    evolves in a non-linear way over time and with growing experience that the
    person has with the robot. For instance, a more experienced user will
    perceive the robot as more familiar and predictable, and in turn
    anthropomorphizes the robot less: \textit{anthropomorphism as a dynamic
    phenomenon}.

\subsection{Article overview}

\begin{inparaenum}[\itshape 1)\upshape]

The remainder of this article is organized as follows.  First, in Section
\ref{sec:overview-studies}, we give a brief overview on the various studies that
we conducted to investigate anthropomorphism, namely \item the Roomba study,
\item the Forum study, \item the Ranger domino study. Next, Section
\ref{sec:anthropomorphism} aims to establish a common ground in the
understanding of anthropomorphism. We first explore an artifact-centered
perspective and briefly outline how the human-like design of agents can
encourage anthropomorphism.  Then, we explore a human-centered perspective
by integrating theories from cognitive science, developmental and social
psychology that serve to explain anthropomorphism. At the end of this
section we bring both perspectives together and interpret that
anthropomorphism is a multi-factor phenomenon.  Based on the two previous
sections, Section~\ref{sec:our-ideas} introduces an initial model of the
\textit{dynamics of anthropomorphism}. This model describes the variances
and dynamics of people's tendency to anthropomorphize an artificial agent,
such as a robot.  Then Section \ref{sec:cognition-neuroscience} presents a
cognitive and neuroscience perspective on anthropomorphism to consolidate
our considerations and the proposed model.  This is followed by a discussion
about the limitations of our approach (Section \ref{sec:discussion}) and the
existing and alternative ways to measure anthropomorphism in human-robot
interactions.  Finally, Section \ref{sec:conclusion} concludes this article
and offers starting points for further research on anthropomorphism and the
effect of the human-like design approach in robotics.

\end{inparaenum}
